\texttt{GLAD-M25} - global adjoint tomography model

### 2018-09-28 main.tex
Updates from ShareLaTeX
update depth_slice figures
update till section: Comparison of Histograms

%%gji_extra_guide.tex

@@

\usepackage{graphicx}
\graphicspath{ {./figures} }
\usepackage{subfig}
\usepackage{caption}
\usepackage{floatrow}
\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
\floatsetup[figure]{style=plain,subcapbesideposition=top}
%\usepackage[margin=70mm]{geometry}

@@

\begin{figure}
  \includegraphics[width=0.8\textwidth]{figures/station_map.pdf}
  \caption{Distribution of 11,800 seismographic stations on the globe. Colors denote the number of events which they contribute waveforms to in the model inversion stage. Stations with event response < 400 are plotted in smaller size and they are usually temporary arrays which are deployed over a short period of time.}
  \label{fig:stations}
  \centering
\end{figure}

@@

\begin{equation}
w_{i} = \frac{1}{A} \left(
  \sum_{j=1}^N \exp \left[ \mbox{} - \left( \frac{r_{ij}} {r_{0}} \right)^2 \right] \right) ^{-1}
\label{eq:spatial_weights}
\end{equation}
where $r_{0}$ is the sensitivity distance and $A$ is the normalization term.
The reference distance will control ratios of weightings between dense and
sparse regions. For more details for the weightings, please refer to the Appendix XXX.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/adjoint_workflow_6.pdf}
  \caption{Adjoint Tomography Workflow.}
  \label{fig:phase_hist}
\end{figure}

\subsection{Forward simulation}
The forward simulation takes the CMT sources, stations and Earth models as input. It will generate synthetic seismograms and forward wavefield snapshots. We used SPECFEM3D-GLOBE as the software to calculate the synthetic seismograms with GPU acceleration on Titan at the Oak Ridge Nation Lab.

We simulate 120min of seismograms. It takes 384 Tesla K20 GPUs approximately 10 mins to do the simulation on Titan at the Oak Ridge National Lab for each earthquakes. To fully undo the attenuation, there is also forward wavefield saved to the disk. In out current 17sec resolution and 120min simulation, for each earthquake, the forward wavefield is about 1TB. Given the size the 1480 earthquakes, we have 1.5TB of wavefield during about 10 hours of simulation, which is huge volume of I/O happening during the calculation.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/Preprocess_workflow.pdf}
  \caption{Adjoint Tomography Workflow.}
  \label{fig:phase_hist}
\end{figure}

The seismic data processing takes the raw observed and synthetic data as input, and generate s the adjoint sources as the output. The preprocessing includes: 1) signal processing that  filter into different period bands and rotate the seismograms from East and North to Radial and transverse.; 2) window selection that pick the time window on a pair of observed and synthetic seismograms; 3) frequency-dependent measurement that generate travel-time measurements inside the windows; 4) measurements quality checks that throw out bad windows and associated measurements and only keep good ones; 5) calculate the adjoint sources; 6) for each window calculate its weightings that will be applied to the overall misfit function; 7) construct the final adjoint sources by combining the measurements in different period bands.

\subsection{Adjoint Simulation}
The adjoint simulation takes the adjoint sources as input, and calculate the frechet derivatives as output. The computation cost is twice as the forward simulation since it is both calculating the adjoint wavefiled and reconstruct the forward wavefield. Other than the adjoint sources, the adjoint simulation also reads in the saved forward wave.

The Adjoint simulation now takes about 20 mins for 17sec resolution and 120min simulation, which about the twice the time of forwards simulation, since it needs to calculate both a forward and adjoint wavefield at the same time.

\subsection{Post-processing}
The post-processing takes the frechet derivatives as the input and generate the model updates as the output. It contains: 1) summation of single event kernels; 2) smoothing the summed kernels; 3) Preconditions using pseudo-Hessian; 4) Numerical optimization to get search direction; 5) calculate the step length on the search direction; 6) update the model using the search direction and step length.

Conjugated gradient methods was used from M00 to M21. Then L-BFGS was used afterwards. Here we will discuss the L-BFGS methods in details.

\subsubsection{L-BFGS}

Assuming the model is $m_i$ and gradient is $g_i$ at iteration i, we
defined the change of model and gradient:

\begin{equation*}
    \Delta m_i = m_{i+1} - m_i \quad
    \Delta g_i = g_{i+1} - g_i
\end{equation*}

and
\begin{equation*}
     \rho_k = \frac{1}{\Delta g^{T} \Delta m}
\end{equation*}

The L-BFGS iteration could be described below, at iteration k \citep{wright1999numerical}:

\begin{align*}
q = g_i \\
for \quad i = k -1, k-2, ..., k-n \\
    \alpha_i = \rho_i \delta m_{i}^{T} q \\
    q = q - \alpha_i \delta g_i \\
end for\\
r = H_k^{0}q \\
for i = k-n, k-n+1, ..., k-1 \\
    \beta = \rho_i \delta g_i^{T}r \\
    r = r + \delta m_i(\alpha_i - \beta) \\
end for \\
\end{align*}

The search direction for L-BFGS is:
$$d_i = -r$$

The step length is chosen based upon line search along the search direction of the overall misfit function.

@@

\begin{table}[!htb]
  \centering
  %\label{tab:misfit_reduction_M00_M25}
  \begin{tabular}{|c|c|c|c|}
    \hline
     ~          &  Vertical(\%) & Radial(\%) &  Transverse(\%) \\
    \hline
    17--40s                &    31.1    &       37.1 &       42.2 \\
    40--100s body waves    &    31.3    &       40.2 &       42.0 \\
    40--100s surface waves &    50.1    &       51.1 &       53.8 \\
    90--250s               &    36.2    &       38.8 &       38.6 \\
    \hline
  \end{tabular}\\
  \caption{Misfit Reduction from M00 to M25 for 1480 earthquakes used in the inversion}
  \label{table:categories}
\end{table}

\begin{table}[!htb]
  \centering
  %\label{tab:category}
  \begin{tabular}{|c|c|c|c|}
    \hline
    ~          &  Vertical(\%) & Radial(\%) &  Transverse(\%) \\
    \hline
    17--40s                &   17.4   &       21.1 &       24.6 \\
    40--100s body waves    &   16.8   &       21.3 &       21.7 \\
    40--100s surface waves &   28.5   &       28.3 &       28.4 \\
    90--250s               &   14.5   &       13.4 &       25.7 \\
    \hline
  \end{tabular}\\
  \label{table:misfit_reduction_M15_M25}
  \caption{Misfit Reduction from M15 to M25 for 1480 earthquakes used in the inversion}
\end{table}

@@

Also include the P arrival measurements here.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/misfit.pdf}
  \caption{Evolution of phase misfits from GLAD-M15 to GLAD-M25. Each color denotes the different stages of inversions, between which we make changes to the inversion by either tuning the weighting strategies or adding more data. (a) Evolution of the total phase misfit. (b) - (j) Misfit in each period band and component. (e) - {g} From GLAD-M18, we future split the 40-100 sec period band into two measurement categories: body waves and surface waves.}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/globe_vp_S362ANI-LLNL.pdf}
  \caption{Map views of global $V_p$ variations at various depths for our model GLAD-M25(left column), S362ANI+M(middle column) and LLNL-G3Dv3(right column)\citep{simmons2012llnl}. For LLNL-G3Dv3, the depths is labeled on the right bottom based on its own mesh spacing. Other plotting conventions are used similar as in Figure \ref{fig:global-vs}.}
\label{fig:global-vp}
\centering
\end{figure}

@@

>>>>

@@

\usepackage{graphicx}
\graphicspath{ {./figures} }
\usepackage{subfig}
\usepackage{caption}
\usepackage{floatrow}
\usepackage{amssymb}
\usepackage{commath}
\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
\floatsetup[figure]{style=plain,subcapbesideposition=top}
%\usepackage[margin=70mm]{geometry}

@@

\begin{figure}
  \includegraphics[width=0.8\textwidth]{figures/station_map.pdf}
  \caption{Distribution of 11,800 seismographic stations on the globe. Colors denote the number of events which they contribute waveforms to in the model inversion stage. Stations with event response < 400 are plotted in smaller size and they are usually temporary arrays which are deployed over a short period of time. The max event reponse comes from IU.ANMO, which contributeed to 1442 out of 1480 eartquakes in out dataset.}
  \label{fig:stations}
  \centering
\end{figure}

@@

\begin{equation}
w_{i} = \frac{1}{A} \left(
  \sum_{j=1}^N \exp \left[ \mbox{} - \left( \frac{r_{ij}} {r_{0}} \right)^2 \right] \right) ^{-1}
\label{eq:spatial_weights}
\end{equation}
where $r_{0}$ is the sensitivity distance and $A$ is the normalization term.
The reference distance $r_0$ will control ratios of weightings between dense and
sparse regions. For more details for the weightings, please refer to the Appendix XXX.

@@

The flow chart for adjoint tomography is shown in Fig.\ref{fig:adjoint_workflow}.
The workflow starts with a certain number of earthquakes. Here, we used the
inverted CMT sources from the Global CMT catalog. Based on the time
of earthquakes, we downloaded observed data, including time series and StaionXML,
from IRIS and other data centers. The we convert the sources, StationXML and time series
into ASDF files. Each ASDF file usually contains a complete set of data from one
earthquake. The data download and conversion need to be done only once. The raw observed data,
storing in ASDF, could be reused afterwards.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/adjoint_workflow_6.pdf}
  \caption{Adjoint Tomography Workflow.}
  \label{fig:adjoint_workflow}
\end{figure}

\subsection{Forward simulation}
The forward simulation takes the CMT sources, stations and Earth models as input.
We will first generate mesh based on the models. The CMT sources and station will
be fed into the solver to generate synthetic seismograms and forward wavefield snapshots.
SEPCFEM3D\_GLOBE with GPU acceleration is used on Titan at the Oak Ridge Nation Lab for
both mesher and solver part.

To simulate 120min of seismograms at 17sec resolution, it takes 384 Tesla K20X
GPUs approximately 10 mins to run on Titan. To precisly calculate the sensitivity
kernels, there is forward wavefield saved to the disk and those will used in the
adjoint simluation in the later stage. In the current resolution
and simulation length, the forward wavefield saved is above 1 terabytes
for each earthquake. Given the size the 1480 earthquakes, we have 1.5 petabytes of
wavefield saved to disk using ADIOS\citep{liu2014hello} during about 10 hours
of simulation using 15,360 GPUs.
It is huge volume of data saved to distk in a relatively short period of time, in which case
ADIOS really helped to relief the I/O bottleneck.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/Preprocess_workflow.pdf}
  \caption{Preprocessing Workflow for seismic data.}
  \label{fig:preprocess_workflow}
\end{figure}

The seismic data processing takes the raw observed and synthetic data as input,
and generate s the adjoint sources as the output. As shown
in Fig.\ref{fig:preprocess_workflow}, preprocessing includes:
\begin{enumerate}
  \item Signal processing that remove the instrument response from observsed data
    to recover the true ground motion. Both observed and synthetic data will be
    filtered into a certain period band. Then the two horizontal components of the
    sesimograms are roated from North, East to radial and transverse only if the
    StationXML passed the sanity check.
  \item Window selection that picks the time window on a pair of
    observed and synthetic seismograms. Pyflex(python version
    of FLEXWIN\citep{maggi2009automated}), is
    such a tool that automatically generate windows where observed data and
    synthetic data are close enough to make measurements based on user defined
    criteria.
  \item Multi-taper measurement that generates frequency-dependent travel-time
    difference within the windows selected from the last step.
    The output files contains both the measurements and misfit values.
  \item Window cleaning that get rid of bad windows(and associated
    measurements), only keeping good ones based on the statistical analysis of
    measurements.
  \item calculate the adjoint sources with the sorted windows.
  \item For each window and associated measurement, calculate its weightings
    that will be applied at the summation stage and calibrate its contribution
    into the overall misfit function, using the methond we mentioned before.
  \item Construct the final adjoint sources by summing the adjoint source from
    different windows and measurement categories.
\end{enumerate}

Even though it is such a computational demanding process as the forward and adjoint
simulation, only taking about $1\%$ CPU hours in the whole workflow, it is crucial
that directly impact the inversion results. It decides which measurement
will be taken into consideration and have a footprint in the final gradient.
We are very careful at this stage, including a lot of checks, such as the Station and
response integratity check, and also filtering out bad measurements to avoid big outliers.
On the other hand, it is very process-demanding, which means it has complex data and
control flows, that every seismograms will go through quite a few of processes
before the final results is generated. We have millions of seismic traces and ten
millions of windows, it is impossible to human-check every trace and measurements.
I will go into more depth about this part in the following sections, discussing our
efforts how to make the data and process safe and robust.

\subsection{Adjoint Simulation}
The adjoint simulation takes the adjoint sources as input, and calculate the frechet
derivatives as output. The computation cost is about twice as the forward simulation since
it is both calculating the adjoint wavefiled and reconstruct the forward wavefield.
Other than the adjoint sources, the adjoint simulation also reads in the saved forward
wave, to fully incorporate attenuation and more exactly calculate the anelastic
sensitivity kernels.
After the adjoint simulation is done, the save wavefield could be deleted and disk
space would be freed.
One Adjoint simulation takes about 25 mins for 17sec resolution for 120min simulation
at 17s resolution.

\subsection{Post-processing}

The post-processing takes the frechet derivatives as the input and generate
the model updates as the output. It contains:
\begin{enumerate}
  \item Summation of each kernel for each earthquake to obtain one
    gradient for the overall misfit function. The number of kernel
    files will reduce from 1480 to 1 after summing them up. Also,
    there will source weightings applied at this stage, to balance
    the uneven distribution of earthquakes.
  \item Smoothing the misfit gradient using 3D gaussian function, which
    serves as a regularization procedure. Instead of using a changing
    smooth radius based upon the "ray density"\citep{bozdaug2016global},
    we used the fixed value in the same iteration, following\cite{hu2012structure}.
  \item Precondition the smoothed kernels. Following
    \citep{luo2013strategies, zhu2012structure, bozdaug2016global},
    the preconditioner we used is:
    \begin{equation}
      P(\mathbf{x}) = 1 / \int \partial_t^2 \mathbf{s}(\mathbf{x}, t) \cdot \partial_t^2 \mathbf{s}^\dagger(\mathbf{x}, T-t) dt
    \end{equation}
\end{enumerate}

Search direction is calculated based on the misfit gradient from the
post-processsing stage. Search direction is calculated using nonlinear conjugate
gradient methods\citep{tape2007finite} and L-BFGS(starting from GLAD-M22),
Here, we used a slightly modified version of conjugate gradient methods,
adding a restart condition\citep{wright1999numerical} to iterations by checking
if the current and previous gradient are far from orthogonal:

\begin{equation}
  \frac { \abs{g_i \cdot g_{i-1}} }  {\norm{g_{i-1}}^2} \geqslant v
\end{equation}

where the $v$ is set to $0.1$ in out case.

Based upon the misfit gradient obtained from the post-processing stage.
Line search is then used to determine the step length for model update
along the search direction. Forward simulations were run for a subset
of 120 earthquakes with different step length and the one with
the max misfit reudction was picked as the model update step length.

@@

\subsection{Misfit Evolution}

The inversion went through a few stages, mainly due to adding more data
and measurement categories and fine tunning the weighting strategies.

\subsubsection{Stage I: M15 - M17}

In the GLAD-M15, we started with 520 events, mainly to validate the test our
software and workflow, making sure things are working properly.
In GLAD-M16, we increase our dataset to 1040 earthquakes.

For iterations at this stage, seismograms are filtered into 3 period bands,
17-40sec, 40-100sec, 90-250 sec. In the 17-40s, we selected body wave windows. For the
other two, we used all the phases available on the seismograms. Conjugate
gradient method is used to calculate the model update.

The misfit values increase a bit at M16 due to increase of earthquakes.

\subsubsection{Stage II: M18-M21}

The major change at stage II is we split the 40-100 sec period band into two: one with
body wave and the other with only surface wave measurements. Since body waves and
surfaces are sensitive to the different parts of the earth, by splitting these two
phases, we have more controls over spatical distribution of the model update.
Table.\ref{table:measurement_category} shows the measurement categories we used starting
from M18 and we will stick to it till M25.

**Not sure if we need to mention the change of weighting strategy in such a detail here**

Since surface waves are mostly sensitive to shallow
structures and for shallow structures our model performs quite good
already given current dataset, we don't expect to see large misfit
reduction from in those long period surface wave category.
To better bring out the deep structure, We tune
down the weightings of 40-100 sec surface wave, thus reduce their
contribution in the overal misfit function. That explained why we see the
misfit jum(decrease) from M17 to M18 in Fig.\ref{fig:misfit}(h) - (j) and
increase Fig.\ref{fig:misfit}(b)-(d). We are using the same data but
redistribution the contribution from each category by changing the measurements
weight in the overall misfit function.

\begin{table}[!htb]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ~          &  Vertical(\%) & Radial(\%) &  Transverse(\%) \\
    \hline
    17--40~s   &   P-SV body waves          & P-SV body waves           & SH body waves   \\
    40--100~s  &   P-SV body waves          & P-SV body waves           & SH body waves \\
    40--100~s  &   Rayleigh surface waves   & Rayleigh surface waves    & Love surface waves \\
    90--250~s  &   All waves                & All waves                 & All waves \\
    \hline
  \end{tabular}\\
  \caption{12 measurement categories starting from M18 till M25. Before M15, the 40--100 body and surface waves are combined into one in each component.}
  \label{table:measurement_category}
\end{table}

The misfit values for 17-40sec and 40-100s body wavess are decreasing fast at
the first few iterations, and flattens gradually, which may indicate that we are
approaching the model resolution limits given current dataset.

\subsubsection{Stage III: M22 - M25}

We added an extra set of 440 earthquakes into our inversion database. Thus, the total
number of earthquakes increased from 1,040 to 1480. We also bring back weightings
of 40-100sec surface wave as stage I since the newly added 440 was mostly located
in poorly covered regions in our previous dataset and those surface wave meansurements
will help to better constrain the shallow structures at those regions.

The misit jumps, from M21 to M22, is mostly due to the redistribution of weights
and new earthquakes being added to the dataset. Optimization method is changed
from conjugate gradient to L-BFGS. Overall misift values is steadily reducing,
as well as those in each category.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/misfit.pdf}
  \caption{Evolution of phase misfits from GLAD-M15 to GLAD-M25.
  Each color denotes the different stages of inversions, between
  which we make changes to the inversion by either tuning the
  weighting strategies or adding more data.
  (a) Evolution of the total phase misfit. (b) - (j) Misfit in
  each period band and component. (e) - (g) From GLAD-M18,
  we future split the 40-100 sec period band into two
  measurement categories: body(red dimond) waves and surface waves
  (red square).}
  \label{fig:misfit}
\end{figure}

\subsection{Misfit Assessment}

In the last setion, we discussed how the misfit evolves in iterations. However, since
we changed our dataset and misfit functions(mainly weightings) between different stages,
it might not be clear that how much did we improve our model. To precisely measure
our model performance, we calculate synthetic data using the same source using M00, M15
and M25. Windows and weightings are also identical across different iterations.

Table.\ref{table:misfit_reduction_M15_M25} shows the misfit reduction from M15 to M25
for 12 categories. For all categories, there is a significant misfit reduction. Among
the three components, transverse have the largest misfit recution. For the 4 period bands,
40-100s surface waves have the largest misfit reduction which may indicate the 40-100 sec
surface wave might not have good fit between synthetic and observed data compared to
other period band and our inversion is improving it a lot.

Table.\ref{table:misfit_reduction_M00_M25} shows M00 to M25. It is very similar to
Table.\ref{table:misfit_reduction_M15_M25} but with larger values since it also
counts the model improvements from M00 to M15. The best improved category is
the transverse component in 40-100s surface waves, with misfit reduction at
53.8\%. The minimum misfit reudction, on the other hand, is the vertical component
in 17-40s body waves, with 31.1\% misfit reduction.

\begin{table}[!htb]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ~          &  Vertical(\%) & Radial(\%) &  Transverse(\%) \\
    \hline
    17--40s  body waves    &   17.4   &       21.1 &       24.6 \\
    40--100s body waves    &   16.8   &       21.3 &       21.7 \\
    40--100s surface waves &   28.5   &       28.3 &       28.4 \\
    90--250s               &   14.5   &       13.4 &       25.7 \\
    \hline
  \end{tabular}\\
  \caption{Misfit Reduction from M15 to M25 for 1480 earthquakes used in the inversion}
  \label{table:misfit_reduction_M15_M25}
\end{table}

\begin{table}[!htb]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
     ~          &  Vertical(\%) & Radial(\%) &  Transverse(\%) \\
    \hline
    17--40s body waves     &    31.1    &       37.1 &       42.2 \\
    40--100s body waves    &    31.3    &       40.2 &       42.0 \\
    40--100s surface waves &    50.1    &       51.1 &       53.8 \\
    90--250s               &    36.2    &       38.8 &       38.6 \\
    \hline
  \end{tabular}\\
  \caption{Misfit Reduction from M00 to M25 for 1480 earthquakes used in the inversion}
  \label{table:misfit_reduction_M00_M25}
\end{table}

\subsection{Comparison of Histograms}

@@

**Also include the P arrival measurements here.**

@@

%\begin{figure}
%\includegraphics[width=0.9\textwidth]{figures/depth_slice/globe_vp_S362ANI-LLNL.pdf}
%  \caption{Map views of global $V_p$ variations at various depths for our model GLAD-M25(left column), S362ANI+M(middle column) and LLNL-G3Dv3(right column)\citep{simmons2012llnl}. For LLNL-G3Dv3, the depths is labeled on the right bottom based on its own mesh spacing. Other plotting conventions are used similar as in Figure \ref{fig:global-vs}.}
%\label{fig:global-vp}
%\centering
%\end{figure

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/globe_vp_LLNL-GAP.pdf}
  \caption{Map views of global $V_p$ variations at various depths for our model GLAD-M25(left column), S362ANI+M(middle column) and LLNL-G3Dv3(right column)\citep{simmons2012llnl}. For LLNL-G3Dv3, the depths is labeled on the right bottom based on its own mesh spacing. Other plotting conventions are used similar as in Figure \ref{fig:global-vs}.}
\label{fig:global-vp}
\centering
\end{figure}

@@

\newpage
